---
title: "AMF_bigeograph_data_compilation"
author: "Smriti Pehim Limbu"
date: "2025-04-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

##package upload
```{r}
if (!requireNamespace("devtools", quietly = TRUE))
    install.packages("devtools")
if (!requireNamespace("pacman", quietly = TRUE))
    install.packages("pacman")
if (!requireNamespace("janitor", quietly = TRUE))
    install.packages("janitor")
if (!requireNamespace("geodata", quietly = TRUE))
    install.packages("geodata")
if (!requireNamespace("sp", quietly = TRUE))
    install.packages("sp")
if (!requireNamespace("mycobank", quietly = TRUE))
    devtools::install_github("gzahn/mycobank")

library(pacman)
library(mycobank)
pacman::p_load(tidyverse, mycobank, janitor, geodata, sp)
```

#uploading data

```{r}
df_biogeography <- read_csv("rawdata/AMF_Biogeography_Database_June_2023.csv") #Stürmer dataset
full_database_df <- readRDS("rawdata/full_database_df.RDS") #Zahn dataset
```



#Changing species names of Zahn dataset for merge

```{r}


# Create a named vector of replacements
name_corrections <- c(
  "Claroideoglomus candidum" = "Entrophospora candidum",
  "Cetraspora gilmorei" = "Scutellospora gilmorei",
  "Cetraspora nodosa" = "Scutellospora nodosa",
  "Claroideoglomus claroideum" = "Entrophospora claroideum",
  "Claroideoglomus drummondii" = "Entrophospora drummondii",
  "Claroideoglomus etunicatum" = "Entrophospora etunicatum",
  "Claroideoglomus lamellosum" = "Entrophospora lamellosum",
  "Entrophospora lamellosa" = "Entrophospora lamellosum",
  "Funneliformis caledonius" = "Funneliformis caledonium",
  "Funneliformis coronatus" = "Funneliformis coronatum",
  "Funneliformis geosporum" = "Funneliformis geosporus",
  "Funneliglomus sanmartinense" = "Funneliglomus sanmartinensis",
  "Halonatospora panshihalos" = "Halonatospora pansihalos",
  "Redeckera fulva" = "Redeckera fulvum",
  "Rhizophagus microaggregatum" = "Rhizoglomus microaggregatum",
  "Rhizophagus proliferus" = "Rhizophagus prolifer",
  "Scutellospora cerradensis" = "Dentiscutata cerradensis",
  "Scutellospora heterogama" = "Dentiscutata heterogama"
)

# Apply the replacements, change latitude and longitude spellings, and group by sequences
df_global.am <- full_database_df %>%
  mutate(good.names = recode(Species, !!!name_corrections)) %>%
  dplyr::rename(
    Latitude_Decimal = latitude,
    Longitude_Decimal = longitude
  ) %>%  mutate(site=paste(paper_id, Latitude_Decimal, Longitude_Decimal, sep = "_")) %>% group_by(site)%>% mutate(id = 0 + cur_group_id())%>%  
  ungroup() %>% group_by(id, good.names, Latitude_Decimal, Longitude_Decimal, year_of_sampling, Biome, plants_dominant, target_gene, sequencing_platform, sample_type, primers, marker, marker_size) %>% dplyr::summarize(MAT=mean(MAT, na.rm=TRUE), MAP=mean(MAP, na.rm=TRUE), pH=mean(pH, na.rm=TRUE), abundances=mean(abundances, na.rm=TRUE))

```

#Now editing Stürmer dataset, first changing latitue and longitude to decimals and cleaning east west in coordinates

```{r}
#Adding latitude and longitude to Stürmer dataset
# Function to convert degrees, minutes, seconds to decimal degrees
convert_to_decimal_latitude <- function(Lat_Degrees, Lat_Min, Lat_Sec) {
  Lat_decimal_degrees <- Lat_Degrees + (Lat_Min / 60) + (Lat_Sec / 3600)
  return(Lat_decimal_degrees)
}

convert_to_decimal_longitude <- function(Long_Degrees, Long_Min, Long_Sec) {
  Long_decimal_degrees <- Long_Degrees + (Long_Min / 60) + (Long_Sec / 3600)
  return(Long_decimal_degrees)
}

# Apply the function to data frame, but first create good.names
df_biogeography.good.names <- df_biogeography %>% mutate(good.names=paste(Genus, Species, sep = "_"))

df_biogeography_lat.long <- df_biogeography.good.names %>% mutate(Latitude_Decimal = convert_to_decimal_latitude(Lat_Degrees, Lat_Min, Lat_Sec)) %>% mutate( Longitude_Decimal = convert_to_decimal_longitude(Long_Degrees, Long_Min, Long_Sec))

#Make directions consistent for lat and long
df_biogeography.a <- df_biogeography_lat.long %>% mutate(North_South=case_when(North_South=="s"~"S", TRUE~North_South)) %>% mutate(North_South=case_when(North_South=="n"~"N", TRUE~North_South)) %>% mutate(North_South=case_when(North_South=="nn"~"N", TRUE~North_South))  %>% mutate(North_South=case_when(North_South=="m"~"N", TRUE~North_South))  %>% mutate(North_South=case_when(North_South=="nm"~"N", TRUE~North_South)) %>% mutate(East_West=case_when(East_West=="e"~"E", TRUE~East_West))%>% mutate(East_West=case_when(East_West=="w"~"W", TRUE~East_West)) %>%  mutate(East_West=case_when(East_West=="n"~"W", TRUE~East_West)) %>% mutate(East_West=case_when(East_West=="ww"~"W", TRUE~East_West))%>% mutate(East_West=case_when(East_West=="2"~"W", TRUE~East_West))%>% mutate(East_West=case_when(East_West=="N"~"W", TRUE~East_West))

df_biogeography_lat.long.decimal <- within(df_biogeography.a, {
             Latitude_Decimal <- ifelse(North_South=="N", Latitude_Decimal, -Latitude_Decimal)
             Longitude_Decimal <- ifelse(East_West=="E", Longitude_Decimal, -Longitude_Decimal)
})

unique(df_biogeography_lat.long.decimal$good.names)
```

##Editing and cleaning species names in Stürmer dataset for merge

```{r}
name_corrections <- c(
  "Acaulospora_nicolsonii" = "Ambispora_nicolsonii",
  "Acaulospora_paulineae" = "Acaulospora_paulinae",
  "Acaulospora_siverdingii" = "Acaulospora_sieverdingii",
  "Acaulospora_sporocarpica" = "Acaulospora_sporocarpia",
  "Acaulospora_tuberculosa" = "Acaulospora_tuberculata",
  "Acaulospora_undulata" = "Archaeospora_undulata",
  "Ambispora_lepoticha" = "Ambispora_leptoticha",
  "Ambispora_leptotich" = "Ambispora_leptoticha",
  "Archaeospora_ecuadoriana" = "Archaeospora_ecuadoriana",
  "Archaeospora_europaea" = "Archaeospora_europea",
  "Archaeospora_spainiae" = "Archaeospora_spainii",
  "Cetraspora_armeniaca" = "Scutellospora_armeniaca",
  "Cetraspora_gilmorei" = "Scutellospora_gilmorei",
  "Cetraspora_heterogama" = "Dentiscutata_heterogama",
  "Cetraspora_nodosa" = "Scutellospora_nodosa",
  "Cetraspora_pellucida" = "Scutellospora_pellucida",
  "Cetraspora_savannicola" = "Dentiscutata_savannicola",
  "Cetraspora_spinosissima" = "Scutellospora_spinosissima",
  "Claroideoglomus_candidum" = "Entrophospora_candidum",
  "Claroideoglomus_claoroideum" = "Entrophospora_claroideum",
  "Claroideoglomus_claroideum?" = "Entrophospora_claroideum",
  "Claroideoglomus_drummondii" = "Entrophospora_drummondii",
  "Claroideoglomus_etunicatum" = "Entrophospora_etunicatum",
  "Claroideoglomus_hanlinii" = "Entrophospora_hanlinii",
  "Claroideoglomus_lamellosum" = "Entrophospora_lamellosum",
  "Claroideoglomus_luteum" = "Entrophospora_luteum",
  "Claroideoglomus_walkeri" = "Entrophospora_walkeri",
  "Corymbiglomus_corymbiformis" = "Corymbiglomus_corymbiforme",
  "Corymbiglomus_globifera" = "Corymbiglomus_globiferum",
  "Dentiscutata_biornata" = "Scutellospora_biornata",
  "Dentiscutata_erythropa" = "Dentiscutata_erythropus",
  "Dentiscutata_hawaiiensis" = "Scutellospora_hawaiiensis",
  "Dentiscutata_scutata" = "Scutellospora_scutata",
  "Diversispora_aurantium" = "Diversispora_aurantia",
  "Diversispora_iranica" = "Dominikia_iranica",
  "Diversispora_trimulares" = "Diversispora_trimurales",
  "Diversispora_versiforme" = "Glomus_versiforme",
  "Diversispora_versiformis" = "Glomus_versiforme",
  "Dominikia _iranica" = "Dominikia_iranica",
  "Dominikia _minuta" = "Dominikia_minuta",
  "Domnikia_aurea" = "Dominikia_aurea",
  "Entrophospora _infrequens" = "Entrophospora_infrequens",
  "Entrophospora_Infrequens" = "Entrophospora_infrequens",
  "Entrophspora_infrequens" = "Entrophospora_infrequens",
  "Funneliformis_caldeonius" = "Funneliformis_caledonium",
  "Funneliformis_caledonium?" = "Funneliformis_caledonium",
  "Funneliformis_caledonius" = "Funneliformis_caledonium",
  "Funneliformis_coronatus" = "Funneliformis_coronatum",
  "Funneliformis_dimorphicum" = "Funneliformis_dimorphicus",
  "Funneliformis_fragilistratum" = "Funneliformis_fragilistratus",
  "Funneliformis_geosporum" = "Funneliformis_geosporus",
  "Funneliformis_geosporum?" = "Funneliformis_geosporus",
  "Funneliformis_halonatum" = "Funneliformis_halonatus",
  "Funneliformis_monosporum" = "Funneliformis_monosporus",
  "Funneliformis_multiforum" = "Funneliformis_multiforus",
  "Funneliformis_verruculosum" = "Funneliformis_verruculosus",
  "Fuscutata_aurea" = "Dentiscutata_aurea",
  "Glomus _microaggregatum" = "Rhizoglomus_microaggregatum",
  "Glomus_aggreatum" = "Rhizophagus_aggregatus",
  "Glomus_aggregatum" = "Rhizophagus_aggregatus",
  "Glomus_antarcticum" = "Rhizophagus_antarcticus",
  "Glomus_aureum" = "Dominikia_aurea",
  "Glomus_avelingiae" = "Redeckera_avelingiae",
  "Glomus_badium" = "Funneliformis_badium",
  "Glomus_canadense" = "Redeckera_canadensis",
  "Glomus_candidum" = "Entrophospora_candidum",
  "Glomus_delihense" = "Glomus_delhiense",
  "Glomus_epigaeum" = "Diversispora_epigaea",
  "Glomus_fragile" = "Redeckera_fragilis",
  "Glomus_geosporum" = "Funneliformis_geosporus",
  "Glomus_gibbosum" = "Diversispora_gibbosa",
  "Glomus_hederabadensis" = "Glomus_hyderabadensis",
  "Glomus_hyderabadense" = "Glomus_hyderabadensis",
  "Glomus_insculptum" = "Diversispora_insculpta",
  "Glomus_invermaium" = "Rhizophagus_invermaius",
  "Glomus_micraggregatum" = "Rhizoglomus_microaggregatum",
  "Glomus_microaggregatum" = "Rhizoglomus_microaggregatum",
  "Glomus_multiforum" = "Funneliformis_multiforus",
  "Glomus_pustulatum" = "Diversispora_pustulata",
  "Glomus_tenebrosum?" = "Glomus_tenebrosum",
  "Glomus_tenerum" = "Diversispora_tenera",
  "Innospora_majewski" = "Innospora_majewskii",
  "innospora_majewskii" = "Innospora_majewskii",
  "Intrornatospora_intraornata" = "Intraornatospora_intraornata",
  "Microkamieskia_perpusilla" = "Microkamienskia_perpusilla",
  "Mikrokamienskia_perpusilla" = "Microkamienskia_perpusilla",
  "Nanoglomus_plukenetia" = "Nanoglomus_plukenetiae",
  "Oehlia_diapahna" = "Oehlia_diaphana",
  "Oehlia_diaphan" = "Oehlia_diaphana",
  "Pacispora_boliviana" = "Paraglomus_bolivianum",
  "Pacispora_coralloidea" = "Pacispora_coralloidae",
  "Pacispora_dominikii" = "Pacispora_scintillans",
  "Pacispora_robiginia" = "Pacispora_robigina",
  "Paraglomus_brazilianum" = "Paraglomus_brasilianum",
  "Paraglomus_majewskii" = "Innospora_majewskii",
  "Paraglomus_occultm" = "Paraglomus_occultum",
  "Paraglomus_occultum?" = "Paraglomus_occultum",
  "Paraglomus_pernabucanum" = "Paraglomus_pernambucanum",
  "Pervetustus _simplex" = "Pervetustus_simplex",
  "Racocetra_alboreosea" = "Racocetra_alborosea",
  "Racocetra_beninenses" = "Racocetra_beninensis",
  "Scutellospora_weresubiae" = "Racocetra_weresubiae",
  "Redeckera_fulva" = "Redeckera_fulvum",
  "Redeckera_megalocarpum" = "Redeckera_megalocarpa",
  "Rhizoglomus _melanum" = "Rhizophagus_melanus",
  "Rhizoglomus_dalpeae" = "Rhizophagus_dalpeae",
  "Rhizoglomus_dunense" = "Rhizophagus_dunensis",
  "Rhizoglomus_maiae" = "Rhizophagus_maiae",
  "Rhizoglomus_natalense" = "Rhizophagus_natalensis",
  "Rhizoglomus_silesianum" = "Rhizophagus_silesianum",
  "Rhizoglomus_variabile" = "Rhizophagus_variabilis",
  "Rhizoglomus_venetianum" = "Rhizophagus_irregularis",
  "Rhizophagus_aggregatum" = "Rhizophagus_aggregatus",
  "Rhizophagus_clarus?" = "Rhizophagus_clarus",
  "Rhizophagus_claurs" = "Rhizophagus_clarus",
  "Rhizophagus_fasciculatum" = "Rhizophagus_fasciculatus",
  "Rhizophagus_Intraradices" = "Rhizophagus_intraradices",
  "Rhizophagus_irregulare" = "Rhizophagus_irregularis",
  "Rhizophagus_neocaledonicus" = "Silvaspora_neocaledonica",
  "Rhizophagus_proliferum" = "Rhizophagus_prolifer",
  "Rhizophagus_proliferus" = "Rhizophagus_prolifer",
  "Rhizophagus_vesiculiferus" = "Rhizophagus_vesiculifer",
  "Sclerocystis_sinuosum" = "Sclerocystis_sinuosa",
  "Scutellospora_auriglobosa" = "Scutellospora_aurigloba",
  "Scutellospora_dipurparescens" = "Scutellospora_dipurpurescens",
  "Scutellospora_dipurpurascens" = "Scutellospora_dipurpurescens",
  "Scutellospora_dipurpurescans" = "Scutellospora_dipurpurescens",
  "Scutellospora_hawaiiensis" = "Scutellospora_hawaiiensis",
  "Dentiscutata_reticulata" = "Scutellospora_reticulata",
  "Scutellospora_spinossisima" = "Scutellospora_spinosissima",
  "Scutellospora_striata" = "Cetraspora_striata",
  "Septoglomus_constrictus" = "Septoglomus_constrictum",
  "Septoglomus_constricum" = "Septoglomus_constrictum",
  "Sieverdingia _tortuosa" = "Sieverdingia_tortuosa",
  "SIeverdingia_tortuosa" = "Sieverdingia_tortuosa"
)


# Apply the replacements
df_biogeography_lat.long.decimal.clean <- df_biogeography_lat.long.decimal %>% mutate(
    good.names = recode(good.names, !!!name_corrections),
    good.names = gsub("_", " ", good.names),
    Year = as.character(Year))

```

#Further cleaning and editing of Stürmer dataset 

```{r}

df_biogeography_lat.long.decimal.clean %>%
  filter(!is.na(Latitude_Decimal)) %>% 
  filter(!is.na(Longitude_Decimal)) %>% mutate(Year=case_when(Year =="105"~"2015", TRUE~Year)) %>%
  mutate(Year=case_when(Year == "199"~ "1991", TRUE~Year)) %>% 
  mutate(Year=case_when(Year=="20042004"~"2004", TRUE~Year)) %>% 
  mutate(Year=case_when(Year=="19891989"~"1989", TRUE~Year)) %>%  names
#mutate id for Stürmer dataset starting from 1430 and select only necessary columns
df_biogeography.final <- df_biogeography_lat.long.decimal.clean %>%
  filter(!is.na(Latitude_Decimal)) %>% 
  filter(!is.na(Longitude_Decimal)) %>% mutate(Year=case_when(Year =="105"~"2015", TRUE~Year)) %>%
  mutate(Year=case_when(Year == "199"~ "1991", TRUE~Year)) %>% 
  mutate(Year=case_when(Year=="20042004"~"2004", TRUE~Year)) %>% 
  mutate(Year=case_when(Year=="19891989"~"1989", TRUE~Year)) %>%  
  mutate(site=paste(`Reference Title`, Latitude_Decimal, Longitude_Decimal, sep = "_")) %>% 
  group_by(site) %>%
  mutate(id = 1429 + cur_group_id()) %>% 
  ungroup() %>%  
  dplyr::select(-c(Order, Family, Genus, Species, `Biogeographical Realm`, `Ecol Region`, Community, Continent, Country, State_Province, Location, Lat_Degrees, Lat_Min, Lat_Sec,North_South, Lat_Origin, Long_Degrees, Long_Min, Long_Sec, East_West, Long_Origin, Hemisphere, `Soil P`, `Soil N`, `Soil C`, `Organic Matter`, Temp_Data, Prec_Origin, Altitude, Altitude_Origin, `Other Literature?`, `Description?`, `Heavy Metals`, Collector, Notes, pH, `Eco region`, site, Year, `Reference Title`))
df_biogeography.final <- janitor::clean_names(df_biogeography.final)

```

#Joining Zahn and Stürmer datasets
```{r}
df_global.am <- janitor::clean_names(df_global.am)
names(df_biogeography.final)
df_metadata<- merge(df_global.am , df_biogeography.final, by=c("good_names", "latitude_decimal", "longitude_decimal", "id"), all = TRUE)

length(unique(df_metadata$good_names))

df_metadata$soil_p_h=as.numeric(df_metadata$soil_p_h)

#edit relevant columns
df_metadata %>% names
df_metadata.correction1 <- 
  df_metadata %>% 
  mutate(biome=coalesce(biome.x, biome.y)) %>% 
  dplyr::select(-c("biome.x", "biome.y")) %>% mutate(sample_type=coalesce(sample_type, source)) %>% dplyr::select(-source) %>% mutate(soil.pH=coalesce(p_h, soil_p_h)) %>% dplyr::select(-c("p_h", "soil_p_h")) %>% mutate(across(c("host"), ~na_if(.x, ""))) 

#edit biome information
df_metadata.final <- df_metadata.correction1%>% mutate(biome = case_when(
    biome == "Tropical and Subtropical Grasslands, Savannas and Shrublands" ~ "Grassland",
    biome == "Temperate Broadleaf and Mixed Forests" ~ "Forest",
    biome == "Tropical and Subtropical Moist Broadleaf Forests" ~ "Forest",
    biome == "Temperate Grasslands, Savannas and Shrublands" ~ "Grassland",
    biome == "Temperate Conifer Forests" ~ "Forest",
    biome == "Mediterranean Forests, Woodlands and Scrub" ~ "Forest",
    biome == "Flooded Grasslands and Savannas" ~ "Grassland",
    biome == "Tropical and Subtropical Dry Broadleaf Forests" ~ "Forest",
    biome == "Deserts and Xeric Shrublands" ~ "Desert",
    biome == "Tropical and Subtropical Coniferous Forests" ~ "Forest",
    biome == "Montane Grasslands and Shrublands" ~ "Grassland",
    biome == "Tundra" ~ "Tundra",
    biome == "Mangroves" ~ "Forest",
    biome == "Boreal Forests/Taiga" ~ "Forest",
    biome == "Mediterranean Forest, Woodlands & Scrubs" ~ "Forest",
    biome == "Temperate Broadleaf & Mixed Forests" ~ "Forest",
    biome == "Boreal Forest/Taiga" ~ "Forest",
    biome == "Desert and Xeric Shrublands" ~ "Desert",
    biome=="woodland"~"Forest",
    biome=="forest"~"Forest",
    biome=="shrubland"~"Shrubland",
    biome=="tundra"~"Tundra",
    biome=="anthropogenic"~"Anthropogenic",
    biome=="aquatic"~"Aquatic",
    biome=="desert"~"Desert",
    biome=="grassland"~"Grassland",
    biome=="cropland"~"Cropland",
    biome=="wetland"~"Wetland",
    TRUE ~ biome
  ))
```

#Extract worldclim data for all lat long data points
# refactored code to use terra package infrastructures
```{r}
# Step 1: Get unique long/lat
df_long.lat <- df_metadata.final %>%
  select(longitude_decimal, latitude_decimal) %>%
  distinct() %>%
  rename(lon = longitude_decimal, lat = latitude_decimal)
points_vect <- vect(df_long.lat, geom = c("lon", "lat"), crs = "EPSG:4326")
# Step 2: Download WorldClim data (mean monthly temperature at 10' resolution)
worldclim_tmean <- worldclim_global(var = "tavg", res = 10, path = tempdir(), version = "2.1")

# Step 3: Convert long/lat dataframe to SpatVector
points.tmean <- vect(df_long.lat, geom = c("lon", "lat"), crs = crs(worldclim_tmean))

# Step 4: Extract temperature values for all months (12 layers)
values.tmean <- extract(worldclim_tmean, points.tmean)

# Step 5: Combine with coordinates (drop ID column from extract output)
df_extracted_from_worldclim.tmean <- cbind(df_long.lat, values.tmean[,-1])

# Step 6: Calculate mean annual temperature
df_worldclim.tmean <- df_extracted_from_worldclim.tmean %>%
  rowwise() %>%
  mutate(mean_annual_temp_worldclim = mean(c_across(starts_with("wc2.1_10m_tavg_")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(lon, lat, mean_annual_temp_worldclim)

# precip
worldclim_prec <- worldclim_global(var = "prec", res = 10, path = tempdir(), version = "2.1")
values.prec <- extract(worldclim_prec, points_vect)
df_worldclim.prec <- cbind(df_long.lat, values.prec[,-1]) %>%
  rowwise() %>%
  mutate(mean_annual_prec_worldclim = mean(c_across(starts_with("wc2.1_10m_prec_")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(lon, lat, mean_annual_prec_worldclim)

# min temp
worldclim_tmin <- worldclim_global(var = "tmin", res = 10, path = tempdir(), version = "2.1")
values.tmin <- extract(worldclim_tmin, points_vect)
df_worldclim.tmin <- cbind(df_long.lat, values.tmin[,-1]) %>%
  rowwise() %>%
  mutate(mean_annual_tmin_worldclim = mean(c_across(starts_with("wc2.1_10m_tmin_")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(lon, lat, mean_annual_tmin_worldclim)

# max temp
worldclim_tmax <- worldclim_global(var = "tmax", res = 10, path = tempdir(), version = "2.1")
values.tmax <- extract(worldclim_tmax, points_vect)
df_worldclim.tmax <- cbind(df_long.lat, values.tmax[,-1]) %>%
  rowwise() %>%
  mutate(mean_annual_tmax_worldclim = mean(c_across(starts_with("wc2.1_10m_tmax_")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(lon, lat, mean_annual_tmax_worldclim)

# wind
worldclim_wind <- worldclim_global(var = "wind", res = 10, path = tempdir(), version = "2.1")
values.wind <- extract(worldclim_wind, points_vect)
df_worldclim.wind <- cbind(df_long.lat, values.wind[,-1]) %>%
  rowwise() %>%
  mutate(mean_annual_wind = mean(c_across(starts_with("wc2.1_10m_wind_")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(lon, lat, mean_annual_wind)

# vapor pressure
worldclim_vapor <- worldclim_global(var = "vapr", res = 10, path = tempdir(), version = "2.1")
values.vapor <- extract(worldclim_vapor, points_vect)
df_worldclim.vapor <- cbind(df_long.lat, values.vapor[,-1]) %>%
  rowwise() %>%
  mutate(mean_annual_vapor_worldclim = mean(c_across(starts_with("wc2.1_10m_vapr_")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(lon, lat, mean_annual_vapor_worldclim)

# bioclim variables
worldclim_bio <- worldclim_global(var = "bio", res = 10, path = tempdir(), version = "2.1")
values.bio <- extract(worldclim_bio, points_vect)
df_worldclim.bio <- cbind(df_long.lat, values.bio[,-1]) %>%
  rename(
    Prec.max.worldclim = wc2.1_10m_bio_16,
    Prec.min.worldclim = wc2.1_10m_bio_17,
    Prec.seasonality.worldclim = wc2.1_10m_bio_15,
    Prec.annual.worldclim = wc2.1_10m_bio_12,
    Temp.seasonality.worldclim = wc2.1_10m_bio_4,
    Temp.annual.worldclim = wc2.1_10m_bio_1,
    Temp.max.worldclim = wc2.1_10m_bio_5,
    Temp.min.worldclim = wc2.1_10m_bio_6
  ) %>%
  select(lon, lat, everything())


#combine the worldclim data files
df_world.clim.all.variables <- df_worldclim.bio %>%
  left_join(df_worldclim.vapor, by = c("lon", "lat")) %>%
  left_join(df_worldclim.wind, by = c("lon", "lat")) %>%
  left_join(df_worldclim.tmax, by = c("lon", "lat")) %>%
  left_join(df_worldclim.tmin, by = c("lon", "lat")) %>%
  left_join(df_worldclim.prec, by = c("lon", "lat"))


```

#Joining worldclim data to the df_metadata.final data

```{r}
df_metadata.final <- df_metadata.final %>% rename(lon=longitude_decimal,lat=latitude_decimal)
df_metadata.final.with.worldclim <- merge(df_metadata.final, df_world.clim.all.variables, by =c("lon", "lat"), all=FALSE) %>% unique()
str(df_metadata.final.with.worldclim)
df_metadata.final.with.worldclim$precipitation=as.numeric(df_metadata.final.with.worldclim$precipitation)

#cleaning it further
unique(df_metadata.final.with.worldclim$sample_type)

df_metadata.final.with.worldclim <- df_metadata.final.with.worldclim %>% mutate(sample_type=case_when(sample_type =="Field spores"~"Field", sample_type == "Field soil"~ "Field", sample_type=="Field sporse"~"Field", sample_type=="Field Soil"~"Field", sample_type=="Field Spores"~"Field", sample_type=="Fields"~"Field",sample_type=="FieldField"~"Field", sample_type=="Fields spores"~"Field", sample_type=="Pot Culture"~"Trap_culture", sample_type=="Pot Cultures"~"Trap_culture", sample_type=="Pot culture"~"Trap_culture", sample_type=="Pot cultures"~"Trap_culture", sample_type=="Trap cultures"~"Trap_culture", sample_type=="Trap culture"~"Trap_culture", sample_type=="Por culture"~"Trap_culture", sample_type=="Molecular data"~"Molecular_data",sample_type=="Molecular Data"~"Molecular_data",sample_type=="Moelcular data"~"Molecular_data", sample_type=="Molecular data "~"Molecular_data", sample_type=="Molecualr data"~"Molecular_data", sample_type=="Y Sfairi"~"Trap_culture",sample_type=="soil"~"Field",sample_type=="topsoil"~"Field",sample_type=="rhizosphere soil"~"Field",sample_type=="root"~"Field",TRUE~sample_type)) 

#Cleaning ppt data from Sturmer dataset
df_metadata.final.clean <- df_metadata.final.with.worldclim %>% mutate(precipitation = case_when(precipitation == 101310~1013.1, precipitation == 171450~1714.5, precipitation == 160970~1609.7, precipitation == 97890~978.9, precipitation == 94000~940, precipitation== 73280~732.8, precipitation== 72000~720, precipitation== 65100~651, precipitation== 60000~600, precipitation== 56200~562, precipitation== 53100~531, precipitation== 51950~519.5, precipitation== 50500~505, precipitation== 49380~493.8, precipitation== 45300~453, precipitation== 45100~451, precipitation== 44000~440, precipitation== 43180~431.8, precipitation== 40600~406, precipitation== 38720~387.2, precipitation== 38600~386, precipitation== 38550~385.5, precipitation== 35000~350, precipitation == 34800~348, precipitation== 31700~317, precipitation== 29600~296, precipitation== 26600~266, precipitation ==26600~266,precipitation== 24200~242, precipitation== 23180~231.8, precipitation== 22250~222.5, precipitation== 20912~209.12, precipitation== 19690~196.9, precipitation== 19370~193.7, precipitation== 16800~168, precipitation== 16755~167.55, precipitation== 16540~165.4, precipitation== 11530~115.3, precipitation== 11530~115.3, precipitation== 9000~90, precipitation== 8000~80, precipitation== 7128~712.8, precipitation== 7000~70, precipitation==6700~67, precipitation== 6132~613.2, precipitation== 6000~60, precipitation==5000~50, precipitation== 4535~453.5, precipitation== 4375~437.5, precipitation== 4300~43,precipitation== 3452~345.2, precipitation== 3128~312.8, precipitation==109000~1090, precipitation==56500~565, TRUE~precipitation)) #cleaning precipitation from Sidney's database

#cleaning temperature data from Sturmer database
df_meta.1 <- df_metadata.final.clean %>% mutate(temperature = case_when(temperature==-1000~-10, temperature==-392~-3.92, temperature==-170~-1.7, temperature==-150~-1.5, temperature== -73~-0.73, temperature==-26~-0.26, temperature==-25~-0.25, temperature==-21~-2.1, temperature==-17~-1.7, temperature==8~0.8, temperature==90~9, temperature==106~10.6, temperature==107~10.7, temperature==110~11, temperature==112~11.2, temperature==113~11.3, temperature==114~11.4, temperature==115~11.5, temperature==118~11.8, temperature==120~1.2, temperature==123~12.3, temperature==125~12.5, temperature==129~12.9, temperature==137~13.7, temperature==138~13.8, temperature==139~13.9, temperature==147~14.7, temperature==150~15, temperature==154~15.4, temperature==155~15.5, temperature==160~16, temperature==161~16.1, temperature==175~17.5, temperature==182~18.2, temperature==183~18.3, temperature==185~18.5, temperature==189~18.9, temperature==190~19, temperature==197~19.7, temperature==199~19.9, temperature==200~20, temperature==202~20.2, temperature==205~20.5, temperature==206~20.6, temperature==210~21, temperature==211~21.1, temperature==212~21.2, temperature==213~21.3, temperature==214~21.4, temperature==217~21.7, temperature==219~21.9, temperature==220~22, temperature==225~22.5, temperature==229~22.9, temperature==230~23, temperature==233~23.3, temperature==235~23.5, temperature==237~23.7, temperature==238~23.8, temperature==240~24, temperature==244~24.4, temperature==249~24.9, temperature==250~25, temperature==255~25.5, temperature==257~25.7, temperature==258~25.8, temperature==260~26, temperature==262~26.2, temperature==270~27, temperature==271~27.1, temperature==272~27.2, temperature==275~27.5, temperature==290~2.9, temperature==300~30, temperature==311~31.1, temperature==400~4.9, temperature==450~4.5, temperature==475~4.75, temperature==490~4.9, temperature==500~5, temperature==540~5.4, temperature==570~5.7, temperature==580~5.8, temperature==600~6, temperature==604~6.04, temperature==610~6.1, temperature==620~6.2, temperature==630~6.3, temperature==650~6.5, temperature==700~7, temperature==710~7.1, temperature==720~7.2, temperature==750~7.5, temperature==800~8, temperature==810~8.1, temperature==820~8.2, temperature==830~8.3, temperature==840~8.4, temperature==850~8.5, temperature==880~8.8, temperature==890~8.9, temperature==900~9, temperature==920~9.2, temperature==923~9.23, temperature==940~9.4, temperature==944~9.44, temperature==950~9.5, temperature==970~9.7, temperature==1000~10, temperature==1100~11, TRUE~temperature))
                                                                      
df_meta <- df_meta.1 %>% mutate(temperature=case_when(temperature==1200~12, temperature==1250~12.5, temperature==1270~12.7, temperature==1320~13.2, temperature==1390~13.9, temperature==1400~14, temperature==1410~14.10, temperature==1450~14.5, temperature==1470~14.7, temperature==1500~15, temperature==1520~15.2, temperature==1550~15.5, temperature==1560~15.6, temperature==1590~15.9, temperature==1620~16.2, temperature==1650~16.5, temperature==1665~16.65, temperature==1670~16.7, temperature==1680~16.8, temperature==1690~16.9, temperature==1700~17, temperature==1730~17.3, temperature==1740~17.4, temperature==1750~17.5, temperature==1760~17.6, temperature==1770~17.7, temperature==1840~18.4, temperature==1850~18.5, temperature==1910~19.1, temperature==1920~19.2, temperature==1930~19.3, temperature==1950~19.5, temperature==1990~19.9, temperature==2020~20.2, temperature==2060~20.6, temperature==2070~20.7, temperature==2120~21.2, temperature==2130~21.3, temperature==2140~21.4, temperature==2150~21.5, temperature==2190~21.9, temperature==2200~22, temperature==2250~22.5, temperature==2260~22.6, temperature==2340~23.4, temperature==2350~23.5, temperature==2390~23.9, temperature==2400~24, temperature==2460~24.6, temperature==2467~24.67, temperature==2490~24.9, temperature==2500~25, temperature==2550~25.5, temperature==2580~25.8, temperature==2600~26, temperature==2610~26.1, temperature==2620~26.2, temperature==2650~26.5, temperature==2660~26.6, temperature==2670~26.7, temperature==2680~26.8, temperature==2710~27.1, temperature==2720~27.2, temperature==2740~27.4, temperature==2750~27.5, temperature==2790~27.9, temperature==2800~28, temperature==2950~29.5, temperature==3000~30, temperature==3150~31.5, temperature==730~7.5, temperature==715~7.15, temperature==170~1.7,temperature==780~7.8,temperature==58~NA, temperature==760~7.6, temperature ==740~7.4, temperature==910~9.1, TRUE~temperature))

unique(df_meta$precipitation)

unique(df_meta$temperature)

#coalesce temperature and precipitation with mat and map
df_meta %>% names
df_meta <- df_meta %>% mutate(map=coalesce(map, precipitation)) %>% dplyr::select(-precipitation) %>% mutate(mat= coalesce(mat, temperature)) %>% dplyr::select(-temperature) %>% filter(biome!="") %>% filter(biome!= "Rock and Ice")

#remove few columns such as traits for the data publication

df_meta_for_data_publication <- df_meta 
length(unique(df_meta_for_data_publication$good_names))#343

```


# clean up species names using mycobank
```{r}
df_meta_for_data_publication %>% names

mb <- get_mycobank_db()
mb_table <- mycobank::get_mycobank_synonyms(df_meta_for_data_publication$good_names,mycobank_db = mb) 
mycobank_lookup <- data.frame(good_names=mb_table$query,current_name=mb_table$current_name)
mycobank_taxonomy <- get_mycobank_taxonomy(mycobank_lookup$current_name,mycobank_db = mb)
mycobank_taxonomy <- data.frame(mycobank_taxonomy = unname(mycobank_taxonomy) %>% str_replace_all(", ",";"),
           current_name = names(mycobank_taxonomy))

df_meta_for_data_publication <- 
  df_meta_for_data_publication %>%
  left_join(mycobank_lookup, by = "good_names") %>%
  mutate(
    current_name = coalesce(current_name, good_names)
  ) %>%
  left_join(mycobank_taxonomy)


write_csv(df_meta_for_data_publication, "df_AMF_biogeography.csv")

```

